\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{titling}
\usepackage{setspace}
\setcounter{secnumdepth}{0} %Table of contents without numbers

\newcommand{\subtitle}[1]{%
  \posttitle{%
    \par\end{center}
    \begin{center}\large#1\end{center}
    \vskip0.5em}%
}
\onehalfspacing

\title{Learning Journal}
\subtitle{FOAR705 2019}
\author{Matthew Clark}
\date{\vspace{-5ex}} %Used to remove date when using \maketitle%
%%%%%%%%%% Title Page & Contents %%%%%%%%%%
\begin{document}
\maketitle
\newpage
\tableofcontents
%%%%%%%%%% New Title %%%%%%%%%%
\newpage
\title{Data Carpentry Journal + Overleaf notes}
\maketitle
%%%%%%%%%% New Entry %%%%%%%%%%
\newpage
\begin{center}
\section{Week 1 Assignment}
Week 1
\end{center}
\noindent
\textbf{Objective:} To attempt to download and restore a backup from 6 months or older\\
\textbf{Action:}
\begin{enumerate}
    \item Identify what file I want to back up
    \begin{itemize}
        \item I will be attempting to restore an old Ableton Live project from 4 years ago from my Google Drive
    \end{itemize}
    \item Open Google Drive, locate the file
    \item Download the file
    \item Open the file in Ableton and see how it works
\end{enumerate}
\textbf{Errors:} No errors were encountered. \\
\textbf{Results:} I was able to successfully download and open a file from 4 years ago. However, it does highlight that my current backup system (Google Drive) will synchronize the most recent item version and replace an older version, meaning I have a backup, but I have no version control within my backup. This will be a point of action for future me to explore down the line.
%%%%%%%%%% New Entry %%%%%%%%%%
\newpage
\begin{center}
\section{Data Carpentry Exercise 1: Formal Data}
Week 2
\end{center}
\noindent
\textbf{Objective:} To work through the first exercise in \textit{01: Format Data} by using examples from \textit{02: Common Mistakes}.
\newline
\textbf{Action:}
\begin{enumerate}
    \item Download the messy data
    \item Open up the data in a spreadsheet program
    \item Analyze what is wrong with the spreadsheet and discuss ways to clean up the the two tabs in order to put them into one spreadsheet
\end{enumerate}
\textbf{Analysis:}
\begin{itemize}
    \item Using multiple data tables within 1 spreadsheet
    \begin{itemize}
        \item Both tabs contain multiple tables. The issue with this is that the computer can correlate columns and rows from different tables to mean the same thing, and if they contain the wrong data, this will end up with a miscorrelation or an error.
        \item \textbf{To solve:} Have all corresponding data written out such that there are no tables and create extra columns for different variables
    \end{itemize}
    \item Using multiple tabs per spreadsheet
    \begin{itemize}
        \item This can cause issues with particular types of data analysis software, and is preferred to avoid this
        \item \textbf{To Solve:} Place all data in one spreadsheet with extra columns for different variable types which correspond to the different countries used.
    \end{itemize}
    \item Putting more than one type of variable per column
    \begin{itemize}
        \item Both tabs have columns which use different data types or variables per column. In the \textit{Mozambique} tab, the livestock count contain the number of livestock, and what type, with no correlation between what the number means in relation to it. There is also a comment in the Water table which states (only in summer), which will confuse analysis software. In the \textit{Tanzania} tab, the livestock table contains columns with both numerical values and boolean values, which causes both confusion for the human reading and the computer analysing the data
        \item \textbf{To Solve:} Create new columns such that each column corresponds to only 1 variable. As for the \textit{Tanzania} data, create 2 columns, one for \textit{animalsObversed} and use a boolean, and a \textit{numOfAnimals} used if the animals were counted, left blank if the data is missing/incomplete as a NULL value.
    \end{itemize}
    \item Zeros and Nulls
    \begin{itemize}
        \item Zeros should be used when there is a 0 observation result, and a NULL should be used when there is missing or incomplete data, and the 2 should not be mixed up. The best representation of NULL is a blank space, as other forms, such as negative numbers, can be misinterpreted as a negative value. The \textit{Mozambique} data contains both negative values, such as '-99' and '-999', while also using blanks and zeros. The \textit{Tanzania} data uses blanks, however due to the amount of blanks used, it is hard to gather whether or not the researcher intended them to be NULLS or zeros
        \item \textbf{To Solve:} Replace negative values as blank spaces, and use zeros to correspond to zero observations. 
    \end{itemize}
    \item Variation in Spelling
    \begin{itemize}
        \item If a computer program is reading the data, it will distinguish a misspelling as a different type of data. For example, in the \textit{Mozambique} data, the researcher misspells 'earth' as 'errth' a couple of times.
        \item \textbf{To Solve:} Clean up data so that all words are spelt correctly
    \end{itemize}
    \item Variations in variable recording
    \begin{itemize}
        \item If a computer program is reading the data, variations in the variable recording could create issues such as errors or different categorization of variables. Most notably is the \textit{Mozambique} 'plots' table, which uses both 'n' and 'no' to mean the same thing, while also using 'yes' and 'yes (only in summer)', which will cause confusion. There is also a number '1', which adds to the confusion.
        \item \textbf{To Solve:} Make sure all variables are denoted the same way; if using a boolean, then do not use numbers as well. Use an extra column to denote if applicable to a particular season.
    \end{itemize}
    \item Using formatting to convey data
    \begin{itemize}
        \item A computer cannot read a coloured cell easily, and if a custom script is designed to accomodate for this, cross-compatibility will be compromised
        \item \textbf{To Solve:} Use a new column to denote a sub-variable instead of a colour to denote a comment
    \end{itemize}
    \item Merging Cells
    \begin{itemize}
        \item The headings used for the tables are located in merged cells. This can cause issues for data analysis software.
        \item \textbf{To Solve:} Do not use merged cells. Rather, create another column which denotes what you are looking at (Eg: typeOfObservation - plot, water, livestock, floorType etc)
    \end{itemize}
    \item Spaces used
    \begin{itemize}
        \item Spaces can cause issues with some programs. It is better to use camelCase of underscores to avoid this issue.
        \item \textbf{To Solve:} Change all cells with spaces between words with camelCase or underscores (eg: floorType, wallType etc)
    \end{itemize}
    \item Special Characters
    \begin{itemize}
        \item Special characters such as parenthesis or asterisks may have a \\
        special meaning in particular programs, causing an unintended input/output. Best to avoid these
        \item \textbf{To Solve:} Treat comments in parenthesis or asterisks as variables and add them into new, different columns.
    \end{itemize}
\end{itemize}
\textbf{Errors:} No errors were involved when downloading or opening the files, or exploring the Data Carpentry website.
\newline
\textbf{Result:} Gained new knowledge about what not to do when recording data in a spreadsheet, and ways to clean up existing data.
\vspace{5mm}
%%%%%%%%%% New Entry %%%%%%%%%%
\newpage
\begin{center}
\section{Data Carpentry Exercise 2: Common Mistakes}
Week 2
\subsection{Metadata}
\end{center}
\textbf{Objective:} To analyse the Clean spreadsheet and discuss useful types of Metadata to improve readability (from \textit{01: Formal Data}.)
\newline
\textbf{Action:}
\begin{enumerate}
    \item Download the clean data
    \item Analyse and record different types of Metadata I would include for this spreadsheet
\end{enumerate}
\textbf{Analysis:}
\begin{itemize}
    \item Some of the keyID's are out of order, any reason to that?
    \item Where abouts are the villages located?
    \item What was the specific questions asked
    \item What is meant by 'muddaub' and 'burntbricks'
    \item Define what is meant by 'membAssoc', 'affectConflict' and 'livCount'
    \item Define what time period 'noMeals' is \\ referring to (Per person, daily, weekly?)
    \item Define what is meant by 'instanceID'
\end{itemize}
\textbf{Errors}: No errors encounted when download or opening the text file, or when analysing the data. The only warning when opening the .csv format is that possible data loss may occur when the workbook is saved in this format. I assume that because the .csv format does not account for formatting, unlike an excel format, that if the data is written out correctly, this is not an issue, however any incorrect formatting could result in a loss of data.\\
\textbf{Results:} In looking into this file, it is good to realise that your data will be read by other people, and having a Metadata text file to explain all the shorthands and the way you gathered your data is crucial for others to understand your data structure. 
\newpage
%%%%%%%%%% New Sub-Entry %%%%%%%%%%
\subsection{Other data from similar fields}
\textbf{Objective:} To explore and analyse other datasets from my field for potential issues in their formatting\\
\textbf{Action:}
\begin{enumerate}
    \item Research potential datasets used in the analysis of modern electronic music
    \item Download, unzip, and open the .csv files in excel
    \item Look through the data set and analyse the fields for any potential issues
\end{enumerate}
\textbf{Errors:} There were no issues that occured when researching, downloading, unzipping, and opening these files. \\
\textbf{Results:}
\begin{itemize}
    \item First data set is titled "DJ Mag Top 100 History Dataset" and was \\ retrieved from https://www.kaggle.com/koki25ando/dj-mag-top-100-\\history-dataset
    \begin{itemize}
        \item  The data set was set up correctly. The only issue was the character ë found in Tiësto's name, and because it is a special character with an umlaut, it gets converted to TiÃ<sto, which also has special characters, and may cause further issues for extra programs
        \item \textbf{To Solve:} Anglicize the name to be Tiesto to avoid any reading issues, and make a note in the Metadata.
    \end{itemize}
    \item Second data set is titled "FMA: A dataset for musical analysis" and was retrieved from https://github.com/mdeff/fma
    \begin{itemize}
        \item Overall, the data set was all laid out correctly. I did spot a question mark in box AJ164 instead of having a blank.
    \end{itemize}
    \item Last data set is titled "Pitchfork's reviews(aka musical's genres battle)" and was retrieved from \\https://www.kaggle.com/bcyphers/pitchfork-reviews/
    \begin{itemize}
        \item I could not find any issue with this data set other than the conversion of special characters under 'content' as it was scrapped off the web. I do like how it has categorised dates as having a column for each variable of the date (One for day, one for month, one for year).
    \end{itemize}
\end{itemize}
%%%%%%%%%% New Entry %%%%%%%%%%
\newpage
\begin{center}
\section{LaTeX in Overleaf}
14 August 2019
\end{center}
\textbf{Objective:} To document my first use of Overleaf, including useful commands to remember later, and errors that were encountered
\newline
\textbf{Action:} Commands that were used so far for future me to remember include:
\begin{itemize}
    \item center: Centre Text
    \item section*: New, unnumbered section
    \item date: Write the specified date
    \item textbf: Bold
    \item textit: Italics
    \item newline: Tells LaTeX to start a new line. Use this with caution.
    \item enumeration: Ordered List. Can be nested.
    \item itemize: Unordered List. Can be nested.
    \item vspace: Creates a vertical space. Useful for formatting. Use mm.
    \item Percentage sign: Comment in code.
    \item newpage: Puts the next text onto a new page
\end{itemize}
\textbf{Errors:} I encountered a few errors so far in using Overleaf:
\begin{itemize}
    \item Formatting code manually (Adding in extra tabs)
    \begin{itemize}
        \item While this didn't break compilation, it did pop up and ask why this was done?
        \item \textbf{Solved by:} Removing any extra tabs in the coding. Error went away
    \end{itemize}
    \item Used the Ampersand symbol in text instead of the word 'and'
    \begin{itemize}
        \item The ampersand symbol did not display as text. The symbol is used as a function in the itemize environment.
        \item \textbf{Solved by:} Replacing it with the word 'and'. You can also use a backslash before the ampersand symbol to avoid it disappearing or causing an error.
    \end{itemize}
    \item Using an underscore
    \begin{itemize}
        \item The underscore cannot be used unless in Math mode.
        \item \textbf{Solved by:} When noting examples, \\ I used camelCase instead. Learned that you can use a backslash before the underscore to use it as plain text.
    \end{itemize}
    \item Overfull hbox
    \begin{itemize}
        \item Found this to occur periodically. According to overleaf, "Overfull hbox messages tell you that some line sticks out over the right margin".
        \item \textbf{Solved by:} There were various manual ways of adjusting hboxes. My quick fix was to use a double backslash as a line break to manual divide the line such that overleaf could not complain that it couldn't fit the entire line in. Will see in future if this quick fix over time breaks the system.
    \end{itemize}
    \item Underfull hbox
    \begin{itemize}
        \item Found that when I incorrectly used the newline function, it would spit out this error. Overleaf states "Underfull hbox messages tell you that some line is poorly typeset (or that you've improperly used newline to leave a vertical space (for example, typing two newline in a row);"
        \item \textbf{Solved by:} checking where I have incorrectly used a double backslash or newline. Error went away
    \end{itemize}
    \item Being able to type in a command as plain text and have it not execute
        \begin{itemize}
            \item If i wanted to type in a command for the previous dot points, it would execute instead of staying as plain text
            \item \textbf{Solved by:} Found out about the 'verbatim' command. Will utilize next time.
        \end{itemize}
    \item Having vspace and newline commands not interact properly, causing the line after to space out
        \begin{itemize}
            \item \textbf{To Solve:} Follow this ordering (vspace first):
            \begin{verbatim}
                \vspace{5mm}
                \newline
            \end{verbatim}
        \end{itemize}
\end{itemize}
\textbf{Results:} Was able to research about errors and ways of formatting LaTeX so that:
\begin{enumerate}
    \item Next time, I can solve errors that I have so far encountered quickly without having to research them
    \item Be able to use the \& and \_ symbols without dropping errors or disappearing
    \item Use the verbatim command to document new commands in full text.
\end{enumerate}
%%%%%%%%%% New Entry %%%%%%%%%%
\newpage
\begin{center}
\section{Scoping Exercise \& Overleaf sync with GitHub}
15 August 2019
\end{center}
\textbf{Objective:}
\begin{enumerate}
    \item To document the use of TeX when completing the scoping exercise within Overleaf
    \item To document committing Overleaf .tex files to GitHub through Overleaf's \textit{sync} function for version control
\end{enumerate}
\textbf{Action:}
\begin{enumerate}
    \item Using TeX in Overleaf
    \begin{itemize}
        \item Create a blank project in Overleaf
        \item Use previous knowledge to complete scoping exercise, document new commands and any new errors
    \end{itemize}
    \item Committing an Overleaf document to GitHub
    \begin{itemize}
        \item Select \textit{Menu -> Sync -> GitHub}
        \item Sign it to GitHub
        \item Create a new repository for the scoping exercise titled \textit{clarkProofOfConcept}
        \item Click "Push Overleaf changes to GitHub
        \item Repeat for the Learning Journal
    \end{itemize}
\end{enumerate}
\textbf{Errors:}
\begin{enumerate}
    \item Using Tex in Overleaf:
    \begin{itemize}
        \item No new errors occurred. 
        \begin{itemize}
            \item Only 2 overfull \verb|\hbox| warning occurred, which was easily corrected with a double slash.
            \item Had to watch the use of \& and \% signs, as I failed to push a slash before the symbol multiple times.
        \end{itemize}
    \end{itemize}
    \item Committing an Overleaf document to GitHub
    \begin{itemize}
        \item No errors occurred when syncing the scoping exercise to a new repository.
        \item I was unable to sync my learning journal to an existing repository, so I deleted the existing repository and re-created it in the sync process. The only solution I could find was through git bash, but my previous knowledge of git has failed me, and thus will look into it next time this issue occurs.
    \end{itemize}
\end{enumerate}
\textbf{Analysis:} New TeX commands that were found and used include:
    \begin{itemize}
        \item Useful packages:
        \begin{itemize}
            \item \verb|\usepackage{titling}|: Useful for creating subtitles
            \item \verb|\usepackage{setspace}|: Used for setting line spacing to normal, 1.5, or double spacing
        \end{itemize}
        \item New Commands:
        \begin{itemize}
            \item \verb|\onehalfspacing|: Set the line spacing to 1.5
            \item \verb|\verb|: Typing out a single command without using a verbatim block
            \item \texttt{Ctrl + B; Ctrl + I; Ctrl + U}: Shortcuts for \verb|\textbf{}| etc.
            \item \verb|texttt{}|: Typewriter font
        \end{itemize}
    \end{itemize}
\textbf{Results:}
\begin{enumerate}
    \item Successfully synchronised both the Learning Journal and Proof of Concept to GitHub
    \item Learned and documented new TeX packages and commands for future me.
\end{enumerate}
%%%%%%%%%% New Entry %%%%%%%%%%
\newpage
\begin{center}
\section{Data Carpentry Exercise 3: Dates as Data}
Week 4
\end{center}
\textbf{Objective:}
\begin{enumerate}
    \item To convert a dd/mm/yyyy column into 3 columns for each date variable
    \item To add a new date entry and analyse the result of the year column
\end{enumerate}
\textbf{Action:}
\begin{enumerate}
    \item Download and open SAFI\_dates.xlsx
    \item Create 3 new columns: day; month; year
    \item Change the formatting of these columns from \textit{dates} to \textit{number}
    \item Use Excel's in-built functions to extract each variable\\ (\verb|=DAY(); =MONTH(); =YEAR();|)
    \item Add the date \textit{17/11} to the end of \textit{interview\_date} and analyse the results
\end{enumerate}
\textbf{Analysis}
\begin{itemize}
    \item Was relatively simple to convert \textit{interview\_date} to its new columns
    \item When I entered the new date without a year, it automatically changed it to the current year (\textit{2019}), which could be an issue if the researcher forgot to enter the year.
\end{itemize}
\textbf{Errors:}
\begin{itemize}
    \item I used the Excel functions before formatting the cell's data type to \textit{number}, and what it would do was display a date with the correct variable I was converting, and setting the rest of the variables to the default.
    \begin{itemize}
        \item Example: \verb|=DAY(17/11/2016)| turns into:\\ \textit{17/01/1900}
    \end{itemize}
    After realizing what I had done and changed the format to \textit{number}, it was good to know how it functioned.
\end{itemize}
\textbf{Result:}
\begin{itemize}
    \item Was able to create 3 new columns per date variable, and analyse a date input with no year attached to it.
\end{itemize}
%%%%%%%%%% New Entry %%%%%%%%%%
\newpage
\begin{center}
\section{Data Carpentry Exercise 4: Quality Assurance}
Week 4
\end{center}
\textbf{Objective:} To utilize Excel's ability to restrict the input of a certain cell to a particular form of data or data range \\
\textbf{Action:}
\begin{enumerate}
    \item Download and open a fresh \verb|SAFI_clean.csv| file
    \item Select an applicable column to restrict its input range, and use Excel's \verb|Data Validation| tool to execute it.
    \item Select an applicable column to restrict its input set to a list, and use Excel's \verb|Data Validation| tool to execute it.
\end{enumerate}
Excel's \verb|Data Validation| is located: \begin{center}
\verb|Data > Data Tools > Data Validation|
\end{center}
\textbf{Errors:}
No errors occurred during this exercise.\\
\textbf{Result:}
\begin{itemize}
    \item I was able to set up a range restriction on \verb|years_lived| to \\\verb|between 0 - 123| alongside an error message and tested it out, which worked as expected.
    \item I was able to set up a list restriction on \verb|memb_assoc| to \verb|yes, no, NULL| alongside an error message and tested it out, which worked as expected
\end{itemize}
%%%%%%%%%% New Entry %%%%%%%%%%
\newpage
\begin{center}
\section{Data Carpentry Exercise 5: Exporting Data}
Week 4
\end{center}
\textbf{Objective:} To save a file in Excel as a \verb|.csv|.\\
\textbf{Action:}
\begin{enumerate}
    \item In excel, go to \verb|File > Save As|
    \item Under \verb|Format|, select \verb|Comma Separated Values|.
    \item Hit Save
    \item Open up the file in Notepad++ and observe the result
\end{enumerate}
\textbf{Analysis:} When opening the document in Notepad++, its easy to observe how a \verb|.csv| functions, and even without special parsing software, you can still understand the general gist of the data presented before you. Its universality means long-time backwards compatibility and minimises potential data loss or data corruption.\\
\textbf{Errors:} No errors were encountered.\\
\textbf{Results:} Was able to export a file as a .csv, and was able to open it in Notepad++.
%%%%%%%%%% New Entry %%%%%%%%%%
\newpage
\begin{center}
\section{Further Analysis of Overleaf}
22 August 2019
\end{center}
\textbf{Objective:} To document further errors encountered when working in Overleaf/LaTeX. \\
\textbf{Errors:} 3 added errors I have encountered thus far include:
\begin{itemize}
    \item The \verb|=| disappears when used as text; \verb|\=| doesn't help.
    \item The use of \verb|_| throws an error (Can only be used in MATH\_MODE; \verb|\_| seems to resolve this error
    \item There was a section where I accidentally used \verb|\(|, which threw an error about MATH\_MODE; removing it fixed the issue.
\end{itemize}
\textbf{Analysis:}
\begin{itemize}
    \item \verb|=| is a function used in MATH\_MODE;
    \item \verb|_| is a function used in MATH\_MODE;
    \item After some research, you can initialize MATH\_MODE by using the \verb|\( \)| braces.
\end{itemize}
\textbf{Results:} Discovered some more about LaTeX's MATH\_MODE, and have \\started using \verb|\verb| more to overcome errors when using special characters as text.
%%%%%%%%%% New Title %%%%%%%%%%
\newpage
\title{Unix Shell Exercises + Overleaf notes}
\maketitle
%%%%%%%%%% New Entry %%%%%%%%%%
\newpage
\section{Unix Shell Exercise 1: Setup \& Introduction}
\textbf{Objective:} To read through and understand the setup and introduction to the Unix Shell\\
\textbf{Action:}
\begin{enumerate}
    \item Open Bash in Ubuntu for Windows, as it was already installed on my computer from my undergraduate degree
    \item Test out the \verb|$ls| command
\end{enumerate}
\textbf{Errors:} No errors were encountered\\
\textbf{Result:} Successfully opened Bash \& tested out the first command
%%%%%%%%%% New Entry %%%%%%%%%%
\newpage
\section{Unix Shell Exercise 2: Navigation}
\textbf{Objective:} To run through the following exercises:
\begin{enumerate}
    \item Exploring More \verb|ls| Flags
    \item Listing Recursively and By Time
    \item Absolute vs. Relative Path 
    \item Relative Path Resolution
    \item \verb|ls| Reading Comprehension
\end{enumerate}
\textbf{Action:}
\begin{enumerate}
    \item Exploring More \verb|ls| Flags: Test out the following flags:
    \begin{enumerate}
        \item \verb|ls -l|
        \item \verb|ls -l -h|
    \end{enumerate}
    \item Listing Recursively and By Time: Test out the following flags:
    \begin{enumerate}
        \item \verb|ls -R|: List recursively
        \item \verb|ls -t|: List by time last changed
        \item \verb|ls -R -t|: Test out what order this is sorted in
    \end{enumerate}
    \item Starting from /Users/amanda/data, which of the following commands could Amanda use to navigate to her hoe directory?
    \item Given the filesystem diagram, if \verb|pwd| displays \verb|/Users/thing|, what will \verb|ls -F ../backup| display?
    \item If \verb|pwd| displays \verb|/Users/Backup|, what command will result in the output \verb|pnas_sub/ pnas_final/ original/|?
\end{enumerate}
\textbf{Errors:}
\begin{enumerate}
    \item Mistook \verb|-l| for \verb|-1| and didn't display what I wanted. Realised and changed the flag. Worked after that.
    \item No errors were encountered.
    \item Was able to navigate my directory around using shell commands without any errors.
    \item Not Applicable
    \item Not Applicable
\end{enumerate}
\textbf{Analysis:}
\begin{enumerate}
    \item When comparing \verb|ls|; \verb|ls -l| \& \verb|ls -l -h|, 
    \begin{enumerate}
        \item \verb|ls -l| displayed the following: \\
        \verb|-rwxrwxrwx 1 root root    13076 Jan 17  2019|
        \item \verb|ls -l -h| displayed the following: \\
        \verb|-rwxrwxrwx 1 root root  13K Jan 17  2019|
    \end{enumerate}
        By the looks of things, it will display the file size and the date last modified. \verb|-h|, according to \textit{Solutions}, stands for \textbf{H}uman Readable, and converts the file size to a more readable format.\\
        It also states the file type \& the links back to the root directory.
    \item: When comparing \verb|ls -R|; \verb|ls -l -t|; \verb|ls -l -R -t|:
    \begin{enumerate}
        \item \verb|ls -R| went through every single file in every single sub-directory in about 10 seconds
        \item \verb|ls -l -t| listed the files/folders in my directory by time last changed, and when used in conjunction with \verb|-l|, I could see the result
        \item \verb|ls -l -R -t|: Listed every single file in every sub-directory by time last changed, presented in long form
    \end{enumerate}
    \item Amanda can use the following:
    \begin{enumerate}
        \item \verb|cd ~|: \verb|~| stands for the home directory, so \verb|cd ~| navigates to the home directory.
        \item \verb|cd ~/data/..|: Would navigate to home, down to data, then back up to home. Over-complicated, but it works.
        \item \verb|cd|: Shortcut to go back home
        \item \verb|cd ..|: Moves back up one level, which is home.
    \end{enumerate}
    \item \verb|ls -F| lists files with their file type; \verb|..| brings the directory up a level; \verb|/backup| navigates the directory down a level to the folder \verb|backup|. Hence, it would display:\\
    \verb|original/ pnas_final/ pnas_sub/|
    \item Because \verb|-r| will list the folders backwards, which is what the given list is listed as, and \verb|-F| will display the file type, which are folders, the correct command would be \verb|ls -r -F|. \verb|ls -r -F /Users/backup| would also work, but is unnecessary as we are already in backup.
\end{enumerate}
\textbf{Results:}
\begin{enumerate}
    \item Was able to test out \verb|ls -l| \& \verb|ls -l -h|.
    \item Was able to test out \verb|ls -R| \& \verb|ls -t|;
    \item Was able to analyse and be able to move directory around using path commands.
    \item Was able to correctly identify the result of the command \verb|ls -F ../backup|
    \item Was able to correctly identify and analyse the correct command.
\end{enumerate}
%%%%%%%%%% New Entry %%%%%%%%%%
\newpage
\section{Unix Shell Exercise 3: Files and Directories}
%
\subsection{Task 1: Creating Files}
%
\textbf{Objective:} Test out the \verb|touch| command and create a blank text file. Analyse what happened, how big is the file, and why would you want to use \verb|touch|?
\textbf{Action:}
Executed the command \verb|touch my_file.txt|\\
\textbf{Errors:} No errors were found\\
\textbf{Analysis:} \verb|touch| creates a new file, in this case a text file, without opening the editor, unlike \verb|nano|. The file is 0 bytes in size. According to the \textit{solution}, this is helpful when programs require a blank file to already be generated so it can output data.\\
\textbf{Results:} Successfully identified what the \verb|touch| command did. Had to use the \textit{Solution} to find out why.
%
\subsection{Task 2: Moving to the Current Folder}
%
\textbf{Objective:} Move 2 files from the current directory to another folder shared by the folder above.\\
\textbf{Action:} Fill in the blanks:\\\verb|$ mv ___/sucrose.dat ___/maltose.dat|\\
\textbf{Errors:} Not applicable\\
\textbf{Analysis:} \verb|mv| will move the located file to the current directory. To locate the files, we must move up a level and then access the old folder \verb|analysed/|. Hence, we execute the command:\\
\verb|mv ../analyzed/sucrose.dat ../analyzed/maltose.dat|.\\
\textbf{Results:} Correctly identified the way to move a file from one folder to another folder using the shell.
%
\subsection{Task 3: Renaming Files}
%
\textbf{Objective:} To identify how to rename a file.\\
\textbf{Action:} Analyse the 4 options and determine which is the correct command.\\
\textbf{Errors:} Not applicable\\
\textbf{Analysis:} To rename a file, we use the \verb|mv| command in the following way:\\\verb|mv statstics.txt statistics.txt|.\\
\textbf{Results:} Correctly identified the right way to rename a file.
%
\subsection{Task 4: Moving and Copying}
%
\textbf{Objective:} To analyse the line-by-line code and identify the result\\
\textbf{Action:} Analyse the code and pick a solution:\\ \verb|proteins-saved.dat recombine|
\textbf{Errors:} The correct answer was \verb|recombine|.\\
\textbf{Analysis:} I forgot that the \verb|..| meant move up a level from where the directory is, not where the file is located, and hence we copied the file above the current directory. Hence, the correct answer was \verb|recombine|.\\
\textbf{Result:} Picked the wrong answer, and analysed the correct answer in \textit{solution}.
%
\subsection{Task 5: Using rm Safely}
%
\textbf{Objective:} To analyse what \verb|-i| does when using the remove command\\
\textbf{Action:} Use the remove command to remove a file:\\\verb|rm -i thesis_backup/quotations.txt|\\
\textbf{Errors:} No errors encountered\\
\textbf{Analysis:} Using the \verb|-i| flag gives us a prompt before removing the file. The shell doesn't have a recycling bin like windows, and hence once a file is removed, it is gone forever (mostly).//
\textbf{Result:} Identify the use of \verb|rm -i|
%
\subsection{Task 6: Copy with Multiple Filenames}
%
\textbf{Objective:} To identify the results when copying multiple files in the same \verb|cp| command\\
\textbf{Action:}
\begin{enumerate}
    \item Explain what happens when the following command is executed:\\
    \verb|$ cp amino-acids.txt animals.txt backup/|
    \item Explain what happens when the following command is executed:\\
    \verb|$ cp amino-acids.txt animals.txt morse.txt |
\end{enumerate}
\textbf{Errors:} No errors were encountered.
\textbf{Analysis:}
\begin{enumerate}
    \item When there are multiple files in the same \verb|cp| command, the last argument is treated as the destination, and the files listed beforehand are copied there. In this case, \verb|amino-acids.txt| \& \texttt{animals.txt} copied to \verb|backup/|.
    \item Because the last argument is treated as the destination, the command would throw an error stating that \texttt{morse.txt} is not a destination.
\end{enumerate}
\textbf{Result:} Learned about copying multiple files and the correct syntax needed to execute \texttt{cp} without throwing an error.
%
\subsection{Task 7: List filenames matching a pattern}
%
\textbf{Objective:} To identify the correct use of wildcards\\
\textbf{Action:} Using the \texttt{molecules/} directory, what command produces the output \texttt{ethane.pdb methane.pdb}?\\
\textbf{Errors:} No errors were encountered.
\textbf{Analysis:} One has to remember that \verb|*| means \textit{zero or more characters}, where as \verb|?| is only a single character wildcard. However, I was able to identify that \verb|ls *t??ne.pdb| was the correct command.\\
\textbf{Result:} Was able to analyse and pick the corrent command for the expected output.
%
\subsection{Task 7: More on wildcards}
%
\textbf{Objective:} To identify the correct use of wildcards for the intended output
\textbf{Action:} Fill in the blanks to correctly output the intended output\\
\textbf{Errors:} No errors were encountered.\\
\textbf{Analysis:} \verb|$ cp *calibration.txt backup/calibration| would copy all calibration files to \verb|backup/calibration|.\\
\verb|$ cp 2015-11-* send_to_bob/all_november_files/| would copy all files \\made in November to \verb|send_to_bob/all_november_files/|.\\
\verb|$ cp *-*-23-dataset* send_to_bob/all_datasets_created_on_a_23rd/|\\ would copy all datastes made on the 23rd to \\\verb|send_to_bob/all_datasets_created_on_a_23rd/|. It could be simplified further to \verb|*-23-dataset*|\\
\textbf{Result:} Was able to correctly identify the correct command for the intended output, however my third answer could be simplified further.
%
\subsection{Task 8: Organising Files and Directories}
%
\textbf{Objective:} To figure out the correct way to move multiple items into a \\subdirectory\\
\textbf{Action:} If \verb|ls -F| resulted in the output:\\ \verb|analyzed/  fructose.dat    raw/   sucrose.dat|, identify the correct \\command to move the \verb|.dat| files into \verb|analyzed/|?\\
\textbf{Errors:} No errors were encountered\\
\textbf{Analysis:} Inferring from the previous tasks, I identified the following command would work:\\
\verb|$ mv fructose.dat sucrose.dat analyzed/|\\
However, this could be further simplified by using a wildcard:\\
\verb|$ mv *.dat analyzed/|\\
\textbf{Result:} Correctly moved 2 files into a subdirectory, however discovered\\ my command could've been simplified by using a wildcard.
%
\subsection{Task 9: Reproduce a folder structure}
%
\textbf{Objective:} To reproduce a folder layout using the \texttt{mkdir} command.\\
\textbf{Action:} From 4 options, pick the correct command(s) which results in the following folder layout:
\begin{verbatim}
    2016-05-20/
    |- data
        |-- processed
        |-- raw
\end{verbatim}
\textbf{Errors:} No errors were encountered\\
\textbf{Analysis:} I correctly picked the 2 correct ways by creating the directory before creating the sub-directories and so forth. I was not sure about the syntax, but you can do the following:
\begin{itemize}
    \item \verb|mkdir 2016-05-20/data| to create the subdirectory \verb|data| if the level above was already created.
    \item \verb|mkdir processed raw| will create 2 folders at the same time in the current directory.
\end{itemize}
\textbf{Results:} Correctly analysed how to create folder structures within the Unix shell.
%%%%%%%%%% New Entry %%%%%%%%%%
\newpage
\section{Syntax \& Commands for the Unix Shell}
\textbf{Objective:} To document helpful syntax and commands learned so far when using the Unix shell.\\
\textbf{Action:}
\begin{itemize}
    \item Shell Syntax
    \begin{itemize}
        \item \verb|/|: Root or Directory
        \item \verb|@|: Link
        \item \verb|*|: Executable
        \item \verb|~|: User's home directory
    \end{itemize}
    \item Shell Commands \& Flags
    \begin{itemize}
        \item \verb|pwd|: Present working directory
        \item \verb|ls|: List sub-directories
        \begin{itemize}
            \item \verb|-a|: Show all files including hidden files
            \item \verb|-F|: Show file type by adding a char to entry
            \item \verb|-l|: Long format: Show extra information
            \item \verb|-R|: List directory tree recursively
            \item \verb|-s|: List by file size
            \item \verb|-t|: List by time \& date
        \end{itemize}
        \item \verb|cd|: Command directory
        \begin{itemize}
            \item \verb|cd <folder>|: Used to move the working directory to that folder
            \item \verb|cd ..|: Move up a level
        \end{itemize}
        \item \verb|mkdir|: Make a directory
        \item \verb|touch|: Generate a new file.
        \item \verb|nano|: Generate a new text file and open it in \textit{nano}
        \item \verb|mv|: Move the located file to the current directory. Can also be used to overwrite/rename.
        \begin{itemize}
            \item \verb|-i|: Asks you for confirmation before moving.
        \end{itemize}
        \item \verb|cp|: Copy the located file to the current directory.
        \begin{itemize}
            \item \verb|-r| Used to copy a directory and all its contents recursively.
        \end{itemize}
        \item \verb|rm|: Remove the located file.
        \begin{itemize}
            \item \verb|-i|: Prompt the user before removing file.
        \end{itemize}
        \item \verb|*|: Zero or more character wildcard.
        \begin{itemize}
            \item \verb|?|: Single char wildcard. Can be used multiple times for multiple char's.
        \end{itemize}
    \end{itemize}
    \item Keyboard Shortcuts
    \begin{itemize}
        \item \verb|Ctrl+L|: Clear the shell window
    \end{itemize}
\end{itemize}
\end{document}
